Wigglyhedra

### 2024-11-07 FPSAC25/wigglyhedraFPSAC.tex
minor corrections

We work over the field \(k = \mathbb{C}\).
Given a quiver \(Q\), we denote by \defn{kQ} its path algebra, which is generated by the idempotent paths \(e_i\) and the arrows \((i,i\pm1)\).
A longer path is denoted \((i_1,\ldots,i_n)\), by the vertices it passes through in order.
\begin{definition}
The \defn{zigzag algebra} \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of \(kQ_m\) by the relations \((i-1, i, i+1) = 0 = (i+1, i, i-1)\) and \((i, i-1, i) = (i, i+1, i)\) for all~\(1 < i < m\).
\end{definition}
The basis elements of \(A_m\) are graded by path-length.
We will consider graded modules over \(A_m\).
Let \(P_i = A_m\cdot e_i\) be the \(i\)th indecomposable projective \(A_m\)-module.
For \(j \in \mathbb{Z}\) and any \(A_m\)-module \(M\), let \(M\langle j \rangle\) denote its graded shift by \(j\).
The triangulated category \(\mathcal{C}_m\) defined below is discussed further in e.g.~\cite[Sect.~2.3.3]{bap.deo.lic:22}.
\begin{definition}
The category \(\mathcal{C}_m\) is the smallest triangulated subcategory of \(D^b(A_m\text{-gr-mod})\) containing all the objects \(P_i\langle j \rangle\).
\end{definition}
Objects in \(\mathcal{C}_m\) are of the form \((M^{\bullet}, d)\), where \(M^{\bullet}\) is a direct sum of graded and homological shifts of objects in \(\{P_1,\ldots, P_m\}\), and \(d \colon M^{\bullet} \to M^{\bullet}\) is a differential that raises homological degree by \(1\).
The \defn{level} of a summand with grading shift \(g\) and homological shift \(h\) is defined to be \(g+h\).
We say that \((M^{\bullet},d)\) is \defn{linear} if each constituent summand has level zero.
We say that an object \((M^{\bullet},d)\) is \defn{spherical} if it has a unique nontrivial endomorphism of graded degree zero, and a unique nontrivial endomorphism of graded degree \(2\), up to scaling.
We refer the reader to~\cite{bel.ber.del:82} for standard results about \(t\)-structures.
\begin{proposition}
  The collection of all linear objects \((M^{\bullet}, d)\) in \(\mathcal{C}_m\) is an abelian subcategory of \(\mathcal{C}_m\), which is the heart of a bounded \(t\)-structure on \(\mathcal{C}_m\).
  Let \(\mathcal{A}_m\) be this abelian subcategory.
\end{proposition}

@@

Now consider \(n+2\) points arranged in a horizontal line.
In addition to the wiggly arcs, we also consider \defn{curves}: these are isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
The \emph{Khovanov--Seidel curve model}~\cite[Sect.~4]{kho.sei:02} is a mapping from curves to objects of \(\mathcal{A}_{n+1}\) up to an overall grading and homological shift, as illustrated in~\cref{fig:curveModel}.
An object in \(\mathcal{A}_{n+1}\) lies in the image of this mapping if and only if it is spherical.
Moreover, the minimal intersection number between any pair of curves is the number of morphisms between the associated objects.

\begin{figure}
\centerline{
	\begin{tabular}{c@{\qquad}c@{\qquad}c}
		Spherical objects in bounded 
		&
		&
		\\
		homotopy category of graded projective 
		&
		$\smash{\xrightarrow[\quad\text{combinatorial model}\quad]{}}$
		&
		Curves
		\\
		modules over the zigzag algebra
		&
		&
                \\
		\begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
			\&           \& P_3 \langle -1 \rangle\arrow{dr} \& \\
			\&           \&                      \& P_2\\
			P_1\langle -5 \rangle\arrow{r}    \& P_1\langle -3 \rangle\arrow{r}   \& P_1 \langle -1 \rangle\arrow{ur} \& 
		\end{tikzcd}
		&
		&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve1}}
		\\[.5cm]
		{\scriptsize cohomology} $\Bigg\downarrow$ {\scriptsize filtration}
		&
		&
		$\Bigg\downarrow$ {\scriptsize cut}
		\\[.5cm]
		Heart of the standard $t$-structure
		&
		$\xrightarrow[\quad\text{combinatorial model}\quad]{}$
		&
		Wiggly arcs
		\\
		\begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
			P_1\langle -1 \rangle \arrow{dr}\& \\
			\& P_2 \\
			P_3 \langle -1 \rangle \arrow{ur}
		\end{tikzcd}
		&
          	&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve2}}
		\\
		\(P_1\)
		&
		&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve3}	}
	\end{tabular}
}
\caption{An illustration of the Khovanov--Seidel curve model and \cref{prop:cohomology-pieces}.}
\label{fig:curveModel}
\end{figure}
\begin{proposition}
\label{prop:cohomology-pieces}
Let \(\alpha\) be any curve on \(n+2\) points, and let \(X\) be the associated object in \(\mathcal{C}_{n+1}\).
\begin{enumerate}
\item The object \(X[j]\) lies in \(\mathcal{A}_m\) for some \(j \in \mathbb{Z}\) if and only if \(\alpha\) is a wiggly arc.
  In this case, \(X \cong H^j(X)[-j]\), and \(H^i(X) = 0\) for \(i \neq j\).
\item If \(\alpha\) is a general curve and \(j \in \mathbb{Z}\), each indecomposable summand of \(H^j(X)\) is a spherical object of \(\mathcal{A}_m\).
\end{enumerate}
\end{proposition}

\begin{proposition}
\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider the set of all indecomposable summands of the nonzero objects \(H^j(X)\).
  Let \(T\) be the corresponding set of wiggly arcs.
  Then
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation,
  \item \(T\) contains at most one of the two irrelevant wiggly arcs.
  \end{enumerate}
\end{proposition}
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\) respectively.
Then the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\Ext_{\mathcal{A}_m}^1(X,X')\).
Thus \(\alpha\) and \(\alpha'\) are compatible (as in~\cref{def:incompatibilityDegree}) if and only if \(\Ext_{\mathcal{A}_m}^1(X, X') = 0\).
\end{proposition}

>>>>

We work over the field \(k = \mathbb{C}\).
Given a quiver \(Q\), we denote by \defn{kQ} its path algebra, which is generated by the idempotent paths \(e_i\) and the arrows \((i,i\pm1)\).
A longer path is denoted \((i_1,\ldots,i_n)\), by the vertices it passes through in order.
\begin{definition}
The \defn{zigzag algebra} \(A_m\) corresponding to the quiver \(Q_m\) depicted in~\cref{fig:am-quiver} is the quotient of the path algebra \(kQ_m\) by the relations \((i-1, i, i+1) = 0 = (i+1, i, i-1)\) and \((i, i-1, i) = (i, i+1, i)\) for all~\(1 < i < m\).
\end{definition}
The basis elements of \(A_m\) are graded by path-length.
We consider graded modules over \(A_m\).
Let \(P_i = A_m\cdot e_i\) be the \(i\)th indecomposable projective \(A_m\)-module.
For \(j \in \mathbb{Z}\) and any \(A_m\)-module \(M\), let \(M\langle j \rangle\) denote its graded shift by \(j\).
The triangulated category~\(\mathcal{C}_m\) defined below is discussed further in e.g.~\cite[Sect.~2.3.3]{bap.deo.lic:22}.
\begin{definition}
The category \(\mathcal{C}_m\) is the smallest triangulated subcategory of \(D^b(A_m\text{-gr-mod})\) containing all the objects \(P_i\langle j \rangle\).
\end{definition}
Objects in \(\mathcal{C}_m\) are of the form \((M^{\bullet}, d)\), where \(M^{\bullet}\) is a direct sum of graded and homological shifts of objects in \(\{P_1,\ldots, P_m\}\), and \(d \colon M^{\bullet} \to M^{\bullet}\) is a differential that raises homological degree by \(1\).
The \defn{level} of a summand with grading shift \(g\) and homological shift \(h\) is defined to be \(g+h\).
We say that \((M^{\bullet},d)\) is \defn{linear} if each constituent summand has level zero.
We say that an object \((M^{\bullet},d)\) is \defn{spherical} if it has a unique nontrivial endomorphism of graded degree zero, and a unique nontrivial endomorphism of graded degree \(2\), up to scaling.
We refer the reader to~\cite{bel.ber.del:82} for standard results about \(t\)-structures.
\begin{proposition}
  The collection of all linear objects \((M^{\bullet}, d)\) in \(\mathcal{C}_m\) is an abelian subcategory of~\(\mathcal{C}_m\), which is the heart of a bounded \(t\)-structure on \(\mathcal{C}_m\).
  Let \(\mathcal{A}_m\) be this abelian subcategory.
\end{proposition}

@@

Now consider \(n+2\) points arranged in a horizontal line.
In addition to the wiggly arcs, we also consider \defn{curves}: these are isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
The \defn{Khovanov--Seidel curve model}~\cite[Sect.~4]{kho.sei:02} is a mapping from curves to objects of \(\mathcal{A}_{n+1}\) up to an overall grading and homological shift, as illustrated in~\cref{fig:curveModel}.
An object in \(\mathcal{A}_{n+1}\) lies in the image of this mapping if and only if it is spherical.
Moreover, the minimal intersection number between any pair of curves is the number of morphisms between the associated objects.

\begin{figure}
\centerline{
	\begin{tabular}{c@{\qquad}c@{\qquad}c}
		Spherical objects in bounded 
		&
		&
		\\
		homotopy category of graded projective 
		&
		$\smash{\xrightarrow[\quad\text{combinatorial model}\quad]{}}$
		&
		Curves
		\\
		modules over the zigzag algebra
		&
		&
                \\[.2cm]
		\begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
			\&           \& P_3 \langle -1 \rangle\arrow{dr} \& \\
			\&           \&                      \& P_2\\
			P_1\langle -5 \rangle\arrow{r}    \& P_1\langle -3 \rangle\arrow{r}   \& P_1 \langle -1 \rangle\arrow{ur} \& 
		\end{tikzcd}
		&
		&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve1}}
		\\[.6cm]
		{\scriptsize cohomology} $\Bigg\downarrow$ {\scriptsize filtration}
		&
		&
		$\Bigg\downarrow$ {\scriptsize cut}
		\\[.6cm]
		Heart of the standard $t$-structure
		&
		$\xrightarrow[\quad\text{combinatorial model}\quad]{}$
		&
		Wiggly arcs
		\\
		\begin{tikzcd}[ampersand replacement=\&, row sep=-0.1cm, column sep=0.5cm]
			P_1\langle -1 \rangle \arrow{dr}\& \\
			\& P_2 \\
			P_3 \langle -1 \rangle \arrow{ur}
		\end{tikzcd}
		&
          	&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve2}}
		\\
		\(P_1\)
		&
		&
		\raisebox{-.5cm}{\includegraphics[scale=1.3]{curve3}	}
	\end{tabular}
}
\caption{An illustration of the Khovanov--Seidel curve model and \cref{prop:cohomology-pieces}.}
\label{fig:curveModel}
\end{figure}
\begin{proposition}
\label{prop:cohomology-pieces}
Let \(\alpha\) be any curve on \(n+2\) points, and let \(X\) be the associated object in \(\mathcal{C}_{n+1}\).
\begin{compactenum}
\item The object \(X[j]\) lies in \(\mathcal{A}_m\) for some \(j \in \mathbb{Z}\) if and only if \(\alpha\) is a wiggly arc.
  In this case, \(X \cong H^j(X)[-j]\), and \(H^i(X) = 0\) for \(i \neq j\).
\item If \(\alpha\) is a general curve and \(j \in \mathbb{Z}\), each indecomposable summand of \(H^j(X)\) is a spherical object of \(\mathcal{A}_m\).
\end{compactenum}
\end{proposition}

\begin{proposition}
\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider the set of all indecomposable summands of the nonzero objects \(H^j(X)\).
  Let \(T\) be the corresponding set of wiggly arcs.
  Then
  \begin{compactenum}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant wiggly arcs.
  \end{compactenum}
\end{proposition}
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\).
Then the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\Ext_{\mathcal{A}_m}^1(X,X')\).
Thus \(\alpha\) and \(\alpha'\) are compatible (as in~\cref{def:incompatibilityDegree}) if and only if \({\Ext_{\mathcal{A}_m}^1(X, X') = 0}\).
\end{proposition}
